Optimal Peak Shaving with Bifacial Solar PV: Simulated Storage Battery Dispatch for Electric Vehicle Charging Load

### 2024-01-10 Paper/paper.tex
sonia feedback

Two new categories of load growth that will strain electric networks in most countries are electric heating and cooling, including traditional air conditioning and heat pumps, and Electric vehicle (EV) charging \cite{Bobmann2015}. Heat pump and air conditioning load factor may be increased with architectural features such as insulation and thermal storage, but building efficiency retrofits often experience barriers such as regulation and financing mechanisms due to permitting and other challenges \cite{Bertone2016}. Meanwhile, extreme weather events, especially heat waves, will likely further reduce the load factor of these loads \cite{Villa2022}. EV charging load differs in that consumers will have the flexibility to slowly charge overnight at home for convenience and to minimize their energy costs. However, workplace, fleet, and public EV charging stations will likely still be required and will experience large peaks as consumers require the convenience of fast charging at stations capable of hundreds of kW per vehicle.

@@

\begin{figure}
  \centering
  \includegraphics[width=7.5cm]{./images/oneline.png}
  \caption{Charging Site One Line Schematic. Simplified one-line schematic of the simulated EV charging station with true measured EV charging power, modeled on-site solar generation, and a modeled stationary battery for economic peak shaving.}
  \label{fig:oneline}
\end{figure}

\begin{equation}
  \label{eq:net-load}
  \begin{split}
    \exists B_d, B_c &\text{ s.t. } T \ge SL \forall t                                    \\      
            & \text{given:}                                                                        \\
    NL     & = L - S                                                                              \\
    SL     & = NL - B_d + B_c                                                                     \\
            & \text{where:}                                                                        \\
    t      & \in \mathbb{R}^+ [h] \text{ is the time step }      \\             
    NL(t)  & \in \mathbb{R} [kW] \text{ is the net load time series }                             \\
    L(t)   & \in \mathbb{R}^+ [kW] \text{ is the EV charging load time series }                   \\
    S(t)   & \in \mathbb{R}^+ [kW] \text{ is the solar production time series }                   \\
    SL(t)  & \in \mathbb{R} [kW] \text{ is the site load time series seen at the electric meter } \\
    B_c(t) & \in \mathbb{R}^+ [kW] \text{ is the battery charging load time series }              \\
    B_d(t) & \in \mathbb{R}^+ [kW] \text{ is the battery discharge time series }                  \\
    T(t)   & \in \mathbb{R}^+ [kW] \text{ is power threshold }                                    \\
  \end{split}
\end{equation}

@@

An optimal peak shaving strategy is used to minimize the total retail electric cost to the EV charging station. The strategy is operational only and assumes the solar production time series and battery capacity are fixed. The state variable controlled by the algorithm is battery charge or discharge power. However, the algorithm reframes the problem as one of choosing a power threshold \(T\ [kW]\) applied to the point of injection to the grid, the electric meter. The battery is then dispatched, within its technical limits, to hold the net load below the threshold. When the algorithm is successful the different thresholds for each TOU period are met for an entire month. If the battery reaches a technical limit and net load exceeds a threshold, the simulation is not necessarily invalid but is not likely to minimize the retail electric cost to the site for that combination of solar and battery size. The peak shaving logic is described in Figure \ref{fig:peakshaving-flowchart}.

@@

\begin{table}[!h]
  \caption{Timeseries Data Summary. The primary data sources are Caltech ACN EV charging power and GOES solar irradiance. Solar production is modeled from irradiance.}
  \label{tab:data-summary}
  \begin{tabularx}{\linewidth}{XXXXX}
    \toprule
    Description                     & Location                                & Type                                                                                                                  & Interval                               & Source                                \\
    \midrule
    EV charging power $(kW_{ac})$   & Jet Propulsion Lab, Pasadena CA, USA    & Measured and aggregated from multiple  EV chargers at single site                                                     & 10 sec (average downsampled to 15 min) & Caltech ACN (site JPL) \cite{Lee2021} \\
    Solar irradiance $(W/m^2)$      & GPS: 34.2013, -118.1721 (2x2 km square) & GOES satellite irradiance                                                                                             & 5 min                                  & NSRDB PSMv3 \cite{Sengupta2018}       \\
    PV array production $(kW_{ac})$ & GPS: 34.2013, -118.1721 (2x2 km square) & Modelled from GOES satellite irradiance, 368 Prism Solar 350 W bifacial modules, 368 Enphase IQH 380 W microinverters & 15 min                                 & SAM \cite{NREL2022}                   \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \caption{Total Retail Electric Cost. Simulated total retail electric cost for three of the case studies (some omitted for conciseness) over the 10 months of available data and a given battery capacity. Cost reduction is calculated with respect to the South 20° base case. The best reduction in total retail cost is \$1422 for the 50\% West 90° case.}
  \label{tab:total-cost}
  \begin{tabularx}{\textwidth}{XXXXXXXX}
    \toprule
    Battery $[kWh]$ & \multicolumn{3}{l}{Cost $[\$]$} & \multicolumn{2}{l}{Cost Reduction $[\$]$} & \multicolumn{2}{l}{Cost Reduction $[\%]$}                                                                   \\
                    & South 20°                       & 50\% West 90°                             & 100\% West 90°                            & 50\% West 90° & 100\% West 90° & 50\% West 90° & 100\% West 90° \\
    \midrule
    25              & 31314                           & 30391                                     & 31730                                     & 923           & -416           & 3.0           & -1.3           \\
    50              & 27132                           & 25773                                     & 27022                                     & 1359          & 110            & 5.0           & 0.4            \\
    75              & 23893                           & 22775                                     & 23625                                     & 1118          & 268            & 4.7           & 1.1            \\
    100             & 21229                           & 19807                                     & 20804                                     & \textbf{1422} & 425            & 6.7           & 2.0            \\
    125             & 18954                           & 17623                                     & 18564                                     & 1331          & 390            & 7.0           & 2.1            \\
    150             & 17094                           & 15886                                     & 17286                                     & 1208          & -192           & 7.1           & -1.1           \\
    200             & 14547                           & 13689                                     & 15282                                     & 858           & -735           & 5.9           & -5.0           \\
    400             & 10730                           & 10969                                     & 12131                                     & -239          & -1401          & -2.2          & -13.1          \\
    600             & 8859                            & 9140                                      & 10333                                     & -281          & -1474          & -3.2          & -16.6          \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \caption{Solar capacity sensitivity analysis based on reduction in total retail cost relative to the base South 20° case. While the 50\% West 90° array continues to achieve the highest cost reduction, the battery capacity increases from 25 kWh to 125 kWh.}
  \label{tab:solar-sensitivity}
  \begin{tabularx}{\textwidth}{XXXX}
    \toprule
    Solar Capacity {[}\(kW_{DC}\){]} & Best Case for Cost Reduction         & Best
    Cost Reduction {[}\${]}          & Best Cost Reduction Battery Capacity
    {[}kWh{]}                                                                            \\
    \midrule
    64.4 (0.5x)                      & 50\% West 90°                        & 260  & 25  \\
    128.8 (1.0x)                     & 50\% West 90°                        & 1422 & 100 \\
    193.2 (1.5x)                     & 50\% West 90°                        & 2211 & 125 \\
    257.6 (2.0x)                     & 75\% West 90°                        & 2579 & 125 \\
    \bottomrule
  \end{tabularx}
\end{table}

>>>>

Two new categories of load growth that will strain electric networks in most countries are electric heating and cooling, including traditional air conditioning and heat pumps, and Electric Vehicle (EV) charging \cite{Bobmann2015}. Heat pump and air conditioning load factor may be increased with architectural features such as insulation and thermal storage, but building efficiency retrofits often experience barriers such as regulation and financing mechanisms due to permitting and other challenges \cite{Bertone2016}. Meanwhile, extreme weather events, especially heat waves, will likely further reduce the load factor of these loads \cite{Villa2022}. EV charging load differs in that consumers will have the flexibility to slowly charge overnight at home for convenience and to minimize their energy costs. However, workplace, fleet, and public EV charging stations will likely still be required and will experience large peaks as consumers require the convenience of fast charging at stations capable of hundreds of kW per vehicle.

@@

\begin{figure}
  \centering
  \includegraphics[width=7.5cm]{./images/oneline.png}
  \caption{Charging Site One Line Schematic. Simplified one-line schematic of the EV charging station with true measured EV charging power, modeled on-site solar generation, and a modeled stationary battery for economic peak shaving.}
  \label{fig:oneline}
\end{figure}

\begin{equation}
  \label{eq:net-load}
  \begin{split}
    \exists B_d, B_c &\text{ s.t. } T \ge SL \forall t                                    \\      
            & \text{given:}                                                                        \\
    NL     & = L - S                                                                              \\
    SL     & = NL - B_d + B_c                                                                     \\
            & \text{where:}                                                                        \\
    t      & \in \mathbb{R}^+ (h) \text{ is the time step }      \\             
    NL(t)  & \in \mathbb{R} (kW) \text{ is the net load time series }                             \\
    L(t)   & \in \mathbb{R}^+ (kW) \text{ is the EV charging load time series }                   \\
    S(t)   & \in \mathbb{R}^+ (kW) \text{ is the solar production time series }                   \\
    SL(t)  & \in \mathbb{R} (kW) \text{ is the site load time series seen at the electric meter } \\
    B_c(t) & \in \mathbb{R}^+ (kW) \text{ is the battery charging load time series }              \\
    B_d(t) & \in \mathbb{R}^+ (kW) \text{ is the battery discharge time series }                  \\
    T(t)   & \in \mathbb{R}^+ (kW) \text{ is power threshold }                                    \\
  \end{split}
\end{equation}

@@

An optimal peak shaving strategy is used to minimize the total retail electric cost to the EV charging station. The strategy is operational only and assumes the solar production time series and battery capacity are fixed. The state variable controlled by the algorithm is battery charge or discharge power. However, the algorithm reframes the problem as one of choosing a power threshold $T (kW)$ applied to the point of injection to the grid, the electric meter. The battery is then dispatched, within its technical limits, to hold the net load below the threshold. When the algorithm is successful the different thresholds for each TOU period are met for an entire month. If the battery reaches a technical limit and net load exceeds a threshold, the simulation is not necessarily invalid but is not likely to minimize the retail electric cost to the site for that combination of solar and battery size. The peak shaving logic is described in Figure \ref{fig:peakshaving-flowchart}.

@@

\begin{table}[!h]
  \caption{Timeseries Data Summary. The primary data sources are Caltech ACN EV charging power and GOES solar irradiance. Solar production is modeled from irradiance.}
  \label{tab:data-summary}
  \begin{tabularx}{\linewidth}{XXXXX}
    \toprule
    Description                     & Location                                & Type                                                                                                                  & Interval                               & Source                                \\
    \midrule
    EV charging power $(kW)$   & Jet Propulsion Lab, Pasadena CA, USA    & Measured and aggregated from multiple  EV chargers at single site                                                     & 10 sec (average downsampled to 15 min) & Caltech ACN (site JPL) \cite{Lee2021} \\
    Solar irradiance $(W/m^2)$      & GPS: 34.2013, -118.1721 (2x2 km square) & GOES satellite irradiance                                                                                             & 5 min                                  & NSRDB PSMv3 \cite{Sengupta2018}       \\
    PV array production $(kW_{ac})$ & GPS: 34.2013, -118.1721 (2x2 km square) & Modelled from GOES satellite irradiance, 368 Prism Solar 350 W bifacial modules, 368 Enphase IQH 380 W microinverters & 15 min                                 & SAM \cite{NREL2022}                   \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \caption{Total Retail Electric Cost. Simulated total retail electric cost for three of the case studies (some omitted for conciseness) over the 10 months of available data and a given battery capacity. Cost reduction is calculated with respect to the South 20° base case. The best reduction in total retail cost is \$1422 for the 50\% West 90° case.}
  \label{tab:total-cost}
  \begin{tabularx}{\textwidth}{XXXXXXXX}
    \toprule
    Battery $(kWh)$ & \multicolumn{3}{l}{Cost $(\$)$} & \multicolumn{2}{l}{Cost Reduction $(\$)$} & \multicolumn{2}{l}{Cost Reduction $(\%)$}                                                                   \\
                    & South 20°                       & 50\% West 90°                             & 100\% West 90°                            & 50\% West 90° & 100\% West 90° & 50\% West 90° & 100\% West 90° \\
    \midrule
    25              & 31314                           & 30391                                     & 31730                                     & 923           & -416           & 3.0           & -1.3           \\
    50              & 27132                           & 25773                                     & 27022                                     & 1359          & 110            & 5.0           & 0.4            \\
    75              & 23893                           & 22775                                     & 23625                                     & 1118          & 268            & 4.7           & 1.1            \\
    100             & 21229                           & 19807                                     & 20804                                     & \textbf{1422} & 425            & 6.7           & 2.0            \\
    125             & 18954                           & 17623                                     & 18564                                     & 1331          & 390            & 7.0           & 2.1            \\
    150             & 17094                           & 15886                                     & 17286                                     & 1208          & -192           & 7.1           & -1.1           \\
    200             & 14547                           & 13689                                     & 15282                                     & 858           & -735           & 5.9           & -5.0           \\
    400             & 10730                           & 10969                                     & 12131                                     & -239          & -1401          & -2.2          & -13.1          \\
    600             & 8859                            & 9140                                      & 10333                                     & -281          & -1474          & -3.2          & -16.6          \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{table}[!h]
  \caption{Solar capacity sensitivity analysis based on reduction in total retail cost relative to the base South 20° case. While the 50\% West 90° array continues to achieve the highest cost reduction, the battery capacity increases from 25 kWh to 125 kWh.}
  \label{tab:solar-sensitivity}
  \begin{tabularx}{\textwidth}{XXXX}
    \toprule
    Solar Capacity $(kW_{DC})$ & Best Case for Cost Reduction         & Best
    Cost Reduction $(\$)$          & Best Cost Reduction Battery Capacity
    $(\$)$                                                                            \\
    \midrule
    64.4 (0.5x)                      & 50\% West 90°                        & 260  & 25  \\
    128.8 (1.0x)                     & 50\% West 90°                        & 1422 & 100 \\
    193.2 (1.5x)                     & 50\% West 90°                        & 2211 & 125 \\
    257.6 (2.0x)                     & 75\% West 90°                        & 2579 & 125 \\
    \bottomrule
  \end{tabularx}
\end{table}
